### 2016-06-29 testbed.tex
.

The \ac{Raspi}s should be setup as similar as possible. One of those
few settings that may be different in the devices is their hostname that
helps the user to distinguish the devices from each other.
%The hostname helps distinguish them from each other.
One approach to create unique hostnames would be to rename the hostname
in each memory card. Another way is to create a start up script to automatically
configure the hostname. We decided to define the hostnames based on the \ac{MAC}
addresses of the \ac{Raspi}s wired ethernet interface.

@@

Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:

@@

%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
%the testbed, it still needs to know which name to give itself.
%We do this using a small Bash script:
We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
hostname in \texttt{nodes.csv}.

@@

Line 1 tells the system to interpret the script using Bash. Line 3 gets the
path of the script itself. Line 4 gets the MAC address. This script will be
located in all \ac{Raspi}s and thus return the MAC address of the node itself
when executed.
Line 4 searches line by line for this mac address.
Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
then it will keep the default hostname of the system.

It is likely that the testbed administrator needs to either change or add
scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
all \ac{Raspi}s. One way around this is to make the \ac{Raspi}s fetch the
configuration scripts during start up before running the configuration scripts.

The below script attempts to fetch the scripts:

@@

# Attempt to fetch new configuration files
if ! wget -O -q --show-progress /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

@@

We have packed the files and uploaded them to a webserver. These can be
retrieved with \texttt{wget} and unzipped into the Raspbian lite image that
we are customizing:

% Get the files
\begin{lstlisting}[]
$ wget http://kom.aau.dk/project/TuneSCode/raspi/rasp_config.zip
$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For the above to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
the http server from which the \ac{Raspi} retrieves the files from.
%"nodes.csv" can be changed using a text editor. For vim, this could be done
%the following way:
%\begin{lstlisting}[]
%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

Finally, to actually make the \ac{Raspi}s change hostname, we have to make
each \ac{Raspi} call the above script when it starts up. Optionally,
also run the update script before running other scripts.
Do this with root permissions by inserting the line below before
\texttt{"exit 0"} using your favorite editor:
%"\$\{ROOTDIR\}/etc/rc.local" using
%your favorite text editor..
%The file
%can be opened with your favorite text editor.

@@

Configuring \texttt{rc.local} makes it really cumbersome for the testbed
administrator to add new scripts since a new script first needs to be added
in \texttt{rasp\_config.zip} and then run by adding an additional line of code
to \texttt{rc.local} in the root image or in all memory cards depending on
the final testbed setup.

A solution to this is to have yet another script in \texttt{rasp\_config.zip}
that calls all the other scripts in the desired order. This would look like
below:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

Now, to use the above script, it is required to change \texttt{rc.local}
slightly:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called through the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at system startup. Had the update script
been called in \texttt{main} instead would mean that changes to the 
\texttt{main} script would also first have been available at next startup.

@@

WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.

All other files could be updated as well.

@@

Provided that qemu-arm is enabled, we should now be able to change root (chroot)
into our Raspbian lite image. Change root is a method i linux that enables
the root (/) to be changed. Thus, enabling a linux installation within a
linux installation.

There are a few commands to be performed before we can change root.
First, to get internet access from within the Raspbian lite image it is needed
to copy our resolv.conf into the filesystem. Line 1 steps into Raspbian lite's
root filesystem. Then line 2 copies resolv.conf.

>>>>

The \ac{Raspi}s should be setup as similar as possible. One of those
few settings that may be different in the devices is their hostname.
%The hostname helps the user to distinguish the devices from each other.
%The hostname helps distinguish them from each other.
One approach to create unique hostnames would be to write the hostname
in each memory card. Another way is to create a start up script to automatically
configure the hostname. We decided to define the hostnames based on the \ac{MAC}
addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
administrator will need to insert the \ac{MAC} addresses and hostnames of
his/her \ac{Raspi}s.

@@

%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:

@@

%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
%the testbed, it still needs to know which name to give itself.
%We do this using a small Bash script:
%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
%hostname in \texttt{nodes.csv}.
A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:

@@

The code lines above do the following:
\begin{itemize}
    \item Line 1: Tells system to interpret the script using \ac{Bash}
    \item Line 3-4: Gets the path to the script itself and the list of hostnames
    \item Line 5: Gets the \ac{MAC} address of the node itself
    \item Line 6: Gets the current hostname
    \item Line 7: Gets the new hostname from the hostname list
    \item Line 10-14: Assigns the new hostname
\end{itemize}

%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
%path of the script itself. Line 4 gets the MAC address. This script will be
%located in all \ac{Raspi}s and thus return the MAC address of the node itself
%when executed.
%Line 4 searches line by line for this mac address.
%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
%then it will keep the default hostname of the system.

It is likely that the testbed administrator needs to either change or add
scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
configuration scripts during start up before running the configuration scripts.

The below script attempts to download/update the scripts:

@@

# Attempt to fetch new configuration files
if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

@@

\begin{itemize}
    \item Line 3-4: Is the url and zip file that should be downloaded
    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
        and prints a warning in case of errors
    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
\end{itemize}

We have packed the files and uploaded them to a webserver. These can be
retrieved with \texttt{wget} and unzipped into the Raspbian lite image:

% Get the files
\begin{lstlisting}[]
$ wget ${URL%/*}/rasp_config.zip
$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
The testbed administrator will therefore need to also edit the \ac{URL} in the
script \texttt{update\_rasp\_config}.
%"nodes.csv" can be changed using a text editor. For vim, this could be done
%the following way:
%\begin{lstlisting}[]
%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

Finally, to actually make the \ac{Raspi}s change hostname, we have to make
each \ac{Raspi} call the above script when it starts up. Optionally,
also run the update script before running other scripts.
Do this with root permissions by inserting the line(s) below before
\texttt{"exit 0"} using your favorite editor:
%"\$\{ROOTDIR\}/etc/rc.local" using
%your favorite text editor..
%The file
%can be opened with your favorite text editor.

@@

Configuring \texttt{rc.local} makes it really cumbersome for the testbed
administrator to add new scripts since it first needs to be included
in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional 
line added to \texttt{rc.local} in the root image or in all memory cards depending on
the final testbed setup.

A solution to this is to make yet another script in \texttt{rasp\_config.zip}
that calls all the other scripts in the desired order. This would look like
below:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

Now, to use the above script, \texttt{rc.local} needs to look like this:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at next system startup.
%Had the update script
%been called in \texttt{main} instead would mean that changes to the 
%\texttt{main} script would also first have been available at next startup.

@@

%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.

%All other files could be updated as well.
It is not only possible to add and configure files to/within the image. One can
also install and update the packages within the image which will be illustrated
in the next section.

@@

Provided that qemu-arm is enabled, we should now be able to change root (chroot)
into our Raspbian lite image. Change root is a method in Linux that enables
the root (/) to be changed. Thus, enabling a Linux installation within a
Linux installation.

There are a few commands to be performed before actually changing root into the
root partion of the image.
First, to get internet access from within the Raspbian lite image it is needed
to copy our \texttt{resolv.conf} into the filesystem. Line 1 steps into Raspbian lite's
root filesystem. Then line 2 copies resolv.conf.
